---
title       : Wealth Projection Simulation
subtitle    : 
author      : Rainer
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : standalone # {standalone, draft}
knit        : slidify::knit2slides
---

## Wealth Projection


<p>We present a simple projection used for financial planning.<p>

<p>Based on the level of risk that you are willing to take and the level of returns that you are expecting, a prediction of your future wealth can be made. </p>

--- .class #id 

## Calculating the expected wealth

<p>The expected wealth can be computed from the initial wealth and the expected returns.</p>

<p>Obviously, the higher the expected returns, the faster the wealth will grow.</p>

```{r echo=FALSE}
calc_ages <- function(age,horizon) {
  ages <- rep(age, horizon)
  for(i in 2:horizon) 
  {
    ages[i] <- ages[i-1]+1
  }
  return(ages)
}

# Calculate the projected wealth given the desired returns
calc_wealth <- function(initWealth, retn, horizon) {
  wealth <- rep(initWealth, horizon)
  retpercent <- retn/100.0
  for(i in 2:horizon)
  {
    wealth[i] <- wealth[i-1]*(1+retpercent)
  }
  return(wealth)
}

w <- calc_wealth(5, 3, 35)
plot(w, type="l", col="blue", xlab="Years", ylab="Projected Wealth (M)", main="Projected Wealth")
```

---- .class #id


## Using the Risk

<p>With higher risks, there will be a greater variability in the actual wealth that is accumulated.</p>

<p>The higher the risks, the greater the spread between the upper and lower lines.</p>

```{r echo=FALSE}
# Calculate the standard deviations given the risks
calc_stds <- function(risk, horizon) {
  stds <- rep(risk, horizon) 
  for(i in 2:horizon)
  {
    stds[i] <- risk*sqrt(i)
  }
  return(stds)
}

# Returns the upper curve which represents the mean + 1 sigma for the
# projected wealth
calc_upper67 <- function(wealth, stds, horizon) {
  upper67 <- rep(wealth[1], horizon) 
  for(i in 2:horizon)
  {
    upper67[i] <- wealth[i]*(1+stds[i]/100.0)
  }
  return(upper67)
}

# Returns the lower curve which represents the mean - 1 sigma for the
# projected wealth
calc_lower67 <- function(wealth, stds, horizon) {
  lower67 <- rep(wealth[1], horizon) 
  for(i in 2:horizon)
  {
    lower67[i] <- wealth[i]*(1-stds[i]/100.0)
  }
  return(lower67)
}

# Calculate the projected wealth given the desired returns
calc_wealth <- function(initWealth, retn, horizon) {
  wealth <- rep(initWealth, horizon)
  retpercent <- retn/100.0
  for(i in 2:horizon)
  {
    wealth[i] <- wealth[i-1]*(1+retpercent)
  }
  return(wealth)
}

w <- calc_wealth(5, 5, 35)
stds <- calc_stds(3, 35)
upper67 <- calc_upper67(w, stds, 35)
lower67 <- calc_lower67(w, stds, 35)

plot(w, type="l", col="blue", xlab="Years", ylab="Projected Wealth (M)", main="Projected Wealth")
lines(upper67, type="l", lty=2, lwd=2, col="green")
lines(lower67, type="l", lty=3, lwd=2, col="forestgreen")

```

---- .class #id


## Putting it together

<p>The graph gives you some sense of how risk and return are related to each other.</p>

<p>However, you will need to assess whether the expected returns and desired risks are appropriate in the current investment climate.</p>

```{r echo=FALSE}
# Calculate the standard deviations given the risks
calc_stds <- function(risk, horizon) {
  stds <- rep(risk, horizon) 
  for(i in 2:horizon)
  {
    stds[i] <- risk*sqrt(i)
  }
  return(stds)
}

# Returns the upper curve which represents the mean + 1 sigma for the
# projected wealth
calc_upper67 <- function(wealth, stds, horizon) {
  upper67 <- rep(wealth[1], horizon) 
  for(i in 2:horizon)
  {
    upper67[i] <- wealth[i]*(1+stds[i]/100.0)
  }
  return(upper67)
}

# Returns the lower curve which represents the mean - 1 sigma for the
# projected wealth
calc_lower67 <- function(wealth, stds, horizon) {
  lower67 <- rep(wealth[1], horizon) 
  for(i in 2:horizon)
  {
    lower67[i] <- wealth[i]*(1-stds[i]/100.0)
  }
  return(lower67)
}

# Calculate the projected wealth given the desired returns
calc_wealth <- function(initWealth, retn, horizon) {
  wealth <- rep(initWealth, horizon)
  retpercent <- retn/100.0
  for(i in 2:horizon)
  {
    wealth[i] <- wealth[i-1]*(1+retpercent)
  }
  return(wealth)
}

w <- calc_wealth(5, 8, 35)
stds <- calc_stds(6, 35)
upper67 <- calc_upper67(w, stds, 35)
lower67 <- calc_lower67(w, stds, 35)

plot(w, type="l", col="blue", xlab="Years", ylab="Projected Wealth (M)", main="Projected Wealth")
lines(upper67, type="l", lty=2, lwd=2, col="green")
lines(lower67, type="l", lty=3, lwd=2, col="forestgreen")
```